Categorical and gradient aspects of wordlikeness\thanks{Thanks to Gene Buckley, Charles Yang, and audiences at NELS 43 for helpful comments.}

### 2012-12-02 1.tex
First complete draft!

This dissertation has two distinct but complementary aims. 
The first is to outline the empirical scope of \emph{phonotactic theory}, the theory of speakers' knowledge of possible and impossible sounds, sound sequences, and words.
The second is to show that the facts of this domain are consonant with a traditional view of phonotactic knowledge as categorical, but inconsistent with an increasingly popular view of phonotactics and phonotactic learning as a type of statistical inference over lexical entries.

If MSCs are epiphenomenal, there are no substantive constraints on underlying representations which are not derived from phonological processes.
This principle---\emph{no static phonotactic constraints}---has interesting ramifications for evaluating certain competing phonological analyses.
Consider Sanskrit aspiration alternations such as \emph{bodhati}-\emph{bhotsyati} `he wakes-he will wake'.
%According to one analysis, which has precedents going back to Pāṇini, the root /budh/ undergoes a process shifting aspiration leftward in certain contexts (e.g., \citeboth{Borowsky1983}, \citeboth{Hoenigswald1965}, \citeboth{Kaye1985}, \citeboth{Sag1974}, \citeyear{Sag1976}, \citeboth{Schindler1976}, \citeboth{Stemberger1980}, \citeboth{Whitney1889}:\S141f.).
%An alternative analysis posits an underlying /bhudh/ and a process of aspirate dissimilation, a synchronic analogue of Grassman's Law (e.g., \citeboth{Anderson1970}, \citeboth{Hoard1975}, \citeboth{Kiparsky1965}:\S3.2, \citeboth{Phelps1973}, \citeboth{Phelps1975b}, \citeboth{Zwicky1965}:109f.).
%Under the latter analysis, multiple surface aspirates (e.g., hypothetical *\emph{bhodhati}) are phonotactically marked; the former account makes no such prediction.
%If the principle of no static phonotactic constraints can be maintained, psycholinguistic tasks could in theory be used to decide between these two accounts.

%While lexical entries are thought to consist of segments, and segments to consist of feature specifications, the phonological form of a word like \emph{spectre} is not ``generated'', in any relevant sense, by the concatenation of segments /s/, /p/, and so on, or by the concatenation of syllables \{spɛk\}$_{\sigma}$ and \{tɚ\}$_{\sigma}$. 
%It would be possible to posit ``lexical concatenation rules'', on analogy with the phrase structure rules of syntax, but it is unclear how the strings produced would be linked with meanings, or how such a system would distinguish between, e.g., \emph{brick} and meaningless \emph{blick}.
%The more appealing alternative is that the phonological form of \emph{spectre}, whatever it is, is an atom of unit of lexical memory. 
%It remains to account for any restrictions on the contents of underlying forms. 
%The theory of \emph{phonotactics} (which is related to \emph{syntax} through the Greek root \emph{táxis} `order') is concerned with these restrictions and how they are related to the mappings between underlying and surface forms for which the phonological component is responsible. 
%\footnote{It is tempting, however, to describe binding principles as ``syntactics''.}

\section{The empirical scope of phonotactic theory}

\section{The architectural basis of phonotactic knowledge}

@@

Early generative phonologists posited that phonotactic ill-formedness derives solely from \emph{morpheme structure constraints}, restrictions on underlying representations \citep[e.g.,][]{Chomsky1966,SPE,SPR,Halle1962}.
These come in two types \citep{Stanley1967}.
\emph{Segment structure constraints} impose restrictions on the underlying segment inventory: e.g., in Russian, voicing is contrastive for all obstruents except /ts, tʃ, x/ \citep[22]{SPR}: [dz, dʒ, ɣ] appear in surface, but not underlying, representations.
\emph{Sequence structure constraints} apply to underlying sequences: an example is given below.

@@

\noindent
This sequence structure constraint specifies the second of a sequence of word-initial consonants as a liquid, so as to preclude underlying /bnɪk/, for example.

However, \citet{Shibatani1973} argues convincingly that not all wordlikeness contrasts can be expressed as constraints on URs.
In German, for instance, obstruent voicing is contrastive, but neutralizes finally: e.g., [ɡʀaːt]-[ɡʀaːtə] `ridge(s)' vs.~[ɡʀaːt]-[ɡʀaːdə] `degree(s)'
.
By hypothesis, the non-alternating `ridge(s)' indicates a root-final /t/, and the [t]-[d] alternation in `degree(s)' a final /d/. 
Since the voicing of root-final obstruents is contrastive, the surface restriction on obstruent voicing is not also a sequence structure constraint.
Yet, \citeauthor{Shibatani1973} claims, German speakers judge voiced obstruent-final nonce words to be ill-formed.\footnote{
    Voicing of final obstruents is also lost in German loanword adaptation: e.g., English \emph{hot dog} becomes [hɑt dɔk] \cite[506]{Ussishkin2003}.}

@@

Whereas \citeauthor{Shibatani1973} argues that morpheme structure constraints are insufficient to account for speakers' knowledge of possible words, other authors suggest that the distinction btween morpheme structure constraints and phonological processes may be artifactual.
\citet[297]{Hale1965}, \citet{Kisseberth1970b}, and \citet[212f.]{Postal1968} all observe the tendency for structural descriptions of phonological processes to reappear among the morpheme structure constraints on the same language (e.g., \citeboth{Hale1965}:297, \citeboth{Kisseberth1970b}, \citeboth{Postal1968}:212f.).
In Russian, for instance, a process of anticipatory assimilation ensures that derived clusters of obstruents agree in voice.\footnote{For sake of discussion, the somewhat more complex behavior of of [v] is ignored here.}

\begin{example}[Russian voice assimilation alternations (adapted from \citealt{SPR}:22f.)]
\label{rovs}
\begin{tabular}{l ll ll}
a. & [ˈʒed.ʒbɨ] & `were one to burn'      & [ˈʒet.ʃl\pal{}i] & `should one burn?'      \\
b. & [ˈmoɡ.bɨ]  & `were (he) getting wet' & [ˈmok.l\pal{}i]  & `was (he) getting wet?' \\
\end{tabular}
\end{example}

\noindent
Similarly, underlying voicing is ``nondistinctive in all but the last member of a cluster of obstruents'' (\citeboth{A74}:283): e.g., [proz\pal{}bə] `request'.
Despite their tantalizing similarity, these two facts are treated as distinct under the traditional view.
This is what 
\citet[?]{KK77} call the \emph{duplication} problem,

In some prior sense, the alternation facts are more privileged.
It is certainly possible to deny that the restrictions on underlying representations are psychologically real since they are ``computationally inert and thus irrelevant to the input-output mapping that the grammar is responsible for'' \citep[18]{PE}.
%Somewhat more to the point, the derivations will come out the same whether or not the grammar includes any statement of the redundancy of obstruent voicing. 
On the other hand, voice assimilation is essential to a concise statement of the surface forms of Russian; the only alternative would result in a huge number of allomorph; the only alternative would be to posit an enormous amount of phonologically sensitive suppletive allomorphy, so that, for instance, /moɡ/ occurs before voiced obstruents and /mok/ elsewhere.
\namecite[205f.]{Dell1973} and \namecite[28f.]{Stampe1973} suggest that the problem is that the distinction between constraints on URs and alternations is unmotivated, and that these different levels are related by a principle now known as \emph{Stampean occultation} \cite[54]{OT}.
In a language like Russian, in which surface obstruent clusters exceptionlessly agree in voicing, there is simply no reason for the language acquisition device to posit underlying hetero-voiced obstruent clusters: obstruent voice assimilation ``occults'' underlying */kb/, for instance.
Were such an underlying form posited, it would surface as [gb] in all contexts.
Unless there is some alternation (e.g., epenthesis) which would reveal /kb/, there is simply no reason for learners to posit this abstract underlying representation.\footnote{
    \emph{Lexicon Optimization} \cite[209]{OT} implements a form of Stampean occultation notable in that it projects all non-alternating surface segments directly into URs.
    For instance, in English, Lexicon Optimization demands underlying /ŋ/ in words like \emph{bank}, even though [ŋ] can be analyzed as an allophone of /n/ before velar consonants (e.g., \citeboth[65f.]{Borowsky1986}:65f.), simplifying the phoneme inventory.
    However, this is not a necessary component of an implementation of \citeauthor{Dell1973} and \citeauthor{Stampe1973}'s insight about the relationship between surface and underlying sequence structure restrictions.
    For instance, the hypothetical /bæŋk/ posited by Lexicon Optimization could be revised to /bænk/, and take a \emph{free ride} (in the sense of \citealt{Zwicky1970}) on the process of nasal place assimilation found elsewhere in English.
    Indeed, this seems desirable, since Lexicon Optimization forces a duplication between underlying and surface constraints: for instance, [ŋ] does not appear word-initially and English speakers have considerable difficulty producing it in this position \cite{Rusaw2009}.
    The allophonic analysis of [ŋ] predicts this fact, since there is no way to derive the [ŋ] allophone in onset position.
    Assuming Lexicon Optimization, the only alternative is to posit an ad-hoc constraint against onset [ŋ] \citep[39]{Jusczyk2002}.
    Future work will consider how purely allophonic relationships could be acquired in a framework like Lexicon Optimization.}
In an architecture like Lexical Phonology, it is even possible to apply a process to individual underlying representations (i.e., at the ``lexical level'') to enforce constraints which are not surface-true.
Consequently, MSCs are otiose, their empirical converage entirely subsumed by phonological processes.

This is similar in principle to the context in which Russian obstruent voicing is discussed by \citet{SPR}.
The principle of biuniquness in vogue at that time distinguishes between, and separates, neutralizing (morphophonemic) and non-neutralizing (phonemic) phonological processes.
In Russian, obstruents participate in voice assimilation whether this neutralizes a phonemic distinction (\ref{rovs}a) or not (\ref{rovs}b): recall that there is no underlying /dʒ/ in Russian.
\citeauthor{SPR} famously argues that biuniqueness (and the distinction between the morphophonemic and phonemic levels that follows from it) entails ``a significant increase in the complexity of the representation'' (24).
While \namecite{Anderson2000} sketches an analysis which preserves biuniqueness without morphophonemic/phonemic duplication, this requires further contested assumptions---contrastive underspecification (against which, see \citeboth{Steriade1995}) and a Duke-of-York derivation, and so is consistent with \citeauthor{SPR}'s claim that biuniquness imposes unnecessary complexities.

Even constraints on underlying representations may be elegantly described in terms of non-contrastive prosodic structures like the syllable \citep[e.g.,][]{Hooper1973,Kahn1976}.
For instance, as \citet{Haugen1956} notes, many restrictions on medial consonant clusters follow from the requirement that word-medial consonant clusters be decomposable into a (possibly null) well-formed coda and a well-formed onset.
Assuming Stampean occultation, syllable structure need not be underlyingly present to derive this constraint, contrary to what has sometimes been claimed (e.g., \citeboth{A74}:255).

If some underlying cluster does not surface faithfully, it

: an underlying representation containing a cluster which could not be decomposed in this fashion would be occulted.

Chapter \ref{gaps} FIXME
Similarly, initial /bn/ in English is occulted by FIXME

\section{Problems for probabilistic phonotactics}

@@

Chapter \ref{gradience} considers the status of intermediate well-formedness ratings in wordlikeness tasks. 
It is argued that intermediate wellformedness ratings are obtained whether or not the categories being rated are graded or not.
A primitive categorical model of wordlikeness using simple prosodic representations is outlined, and shown to predict English speakers' wordlikeness judgements as well as state-of-the-art gradient wellformedness models; a simple model of similarity to existing words is also a reliable predictor of wellformedness ratings.
Once categorical effects are controlled for, current gradient phonotactic models are largely uncorrelated with wellformedness ratings.

Chapter \ref{turkish} considers the relationship between lexical generalizations, phonological alternations, and speakers' nonce word judgements with a focus on Turkish vowel patterns.
It is shown that even exception-filled phonological generalizations may have a robust effect on wellformedness judgements.
However, phonotactic generaliztaions go unlearned when they are not derived from phonological alternations; this fact is independent of ``naturalness''.

Chapter \ref{clusters} investigates medial consonant clusters in English.
Static phonotactic constraints previously proposed by \citet{Pierrehumbert1994} to describe gaps in the inventory of medial clusters are shown to be statistically unsound.
Neutralizing phonological alternations impose robust restrictions on the cluster inventory and the remaining gaps can be attributed to the sparse nature of the lexicon.

% JUNK

@@

%\begin{example}[German \emph{ich}- and \emph{ach-laut}]
%\begin{tabular}{l l l}
%a. & [buːx]   & `book'           \\
%   & [tɔxtər] & `daughter'       \\
%   & [naxt]   & `night'          \\
%b. & [siçt]   & `view'           \\
%   & [ʃpeçt]  & `woodpecker'     \\
%   & [ɡərʏçt] & `rumor'          \\
%   & [knøçəl] & `ankle, knuckle' \\
%   & [flɛçə]  & `surface'        \\
%\end{tabular}
%\end{example}
%
%\emph{Umlaut}, the fronting (and raising) of back vowels in certain morphological contexts, produces the front variant of the dorsal fricative; e.g., [lɔx]-[løçər] `hole-holes', \emph{B}[uːx]-\emph{B}[yːç]\emph{er} `book-books',

>>>>

This dissertation has two distinct but complementary aims. 
The first is to outline the empirical scope of \emph{phonotactic theory}, the theory of speakers' knowledge of possible and impossible sounds, sound sequences, and words.
The second is to show that the core facts of this domain are compatible with a traditional view of phonotactic knowledge as independent of the lexicon, categorical, and closely related to phonological processes, but inconsistent with an increasingly popular view of phonotactics and phonotactic learning as a type of probabilistic inference over the lexicon, and therefore gradient and independent of phonological processes.\footnote{
    Throughout, the term \emph{lexicon} is used in a specific sense of the set underlying representations in some language; this is not meant to imply a position on the possibility that larger, composite linguistic representations are also stored in lexical memory.
    See \citealt{LignosInPressa} for a review of experimental evidence on this question.}

Despite a recent surge of interest in phonotactic theory, the empirical scope of the theory remains poorly defined.
%and consequently, it is not uncommon to find suggestions that speakers do not have any type of phonotactic knowledge at all \citep[e.g.,][17f.]{PE}. 
The first task in developing a theory of phonotactic knowledge, then, is to outline the types of facts that the theory should account for.

\section{The empirical scope of phonotactic theory}
\label{s:espt}

Phonotactic knowledge is evidenced in quite different ways than phonological knowledge.
In Russian, for instance, a process of anticipatory assimilation ensures that derived clusters of obstruents agree in voice.\footnote{
    For sake of discussion, the complex behavior of [v] is ignored here.}

\begin{example}[Russian voice assimilation alternations (adapted from \citealt{SPR}:22f.)]
\label{rovs}
\begin{tabular}{l ll ll}
a. & [ˈʒedʒbɨ] & `were one to burn'      & [ˈʒetʃl\pal{}i] & `should one burn?'      \\
b. & [ˈmoɡbɨ]  & `were (he) getting wet' & [ˈmokl\pal{}i]  & `was (he) getting wet?' \\
\end{tabular}
\end{example}

There is only one concise explanation for the [dʒ]-[tʃ] and [k]-[ɡ] alternations, namely that Russian speakers have internalized a process of voice assimilation.
Now, it is also the case that voicing in underlying representations in Russian is ``nondistinctive in all but the last member of a cluster of obstruents'' \citep[283]{A74}.
It is possible to deny, however, that this restriction on underlying hetero-voiced obstruent clusters is part of the grammatical knowledge of Russian speakers. 
For instance, \citeauthor{PE} write:

\begin{quote}
Even if we, as linguists, find some generalizations in our description of the lexicon, there is no reason to posit these generalizations as part of the speaker's knowledge of their langauge, since they are computationally inert and thus irrelevant to the input-output mappings that the grammar is responsible for. \citep[17f.]{PE}
\end{quote}

\noindent
\citeauthor{PE} go on to label constraints on underlying representations as ``extensional'', and thus irrelevant to generative grammar, in contrast to ``intensional'' statements like phonological rules.
Regarding the facts about Russian obstruents discussed so far, it seems to be clear that the alternation facts are in some sense more privileged.
Were it the case that Russian speakers had not internalized a process of obstruent voicing assimilation, the only alternative explanation for the forms in (\ref{rovs}) is a massive system of phonologically conditioned suppletive allomorphy.
In contrast, were there no contraint on underlying obstruent voicing, however, it is not immediately obvious that \emph{anything} would be different. 
Presumably, this is what \citeauthor{PE} mean when they refer to constraints on underlying representations as ``computationally inert''.
There are also many cases where restrictions on underlying representations hold of underlying representations but not of derived forms.
For instance, hiatus is exceptionally rare in native Turkish roots; of the handful of examples in the Turkish Electronic Living Lexicon \citep{TELL}, many appear to be compounds (e.g., \emph{ısıalan} `endothermic', cf.~\emph{ısı} `heat', \emph{alan} `taker').
Yet, numerous phonological processes give rise to hiatus in derived environments \citep[e.g.,][]{Kabak2007b}.
\citet{Silverman2000} observes that operations like reduplication and truncation also tend to introduce violations of constraints on underlying representations.
However, there are many facts which strongly suggest that speakers internalize phonotactic constraints.

\subsection{Wordlikeness judgements}

As \citet{Halle1962} and \citet{Chomsky1965} note, speakers can distinguish between a well-formed and an ill-formed word, neither of which is an actual word.
Neither [blɪk] \emph{blick} nor [bnɪk] \emph{bnick} is a word of English, yet English speakers immediately report that only the former is ``possible'', an \emph{accidental} gap in the lexicon, whereas the latter is judged to be impossible (i.e., structurally excluded).
There can be no question that it is part of speakers' knowledge.
Elicited in a controlled fashion, these \emph{wordlikeness judgements} are perhaps the most important (and least controversial) source of phonotactic evidence, and they play a major role in chapters \ref{gradience}--\ref{turkish} of this dissertation.

\subsection{Word production and recognition}
\label{ss:wpr}

Speakers have difficulty producing \citep[e.g.,][]{Davidson2006a,Davidson2010,Rose2007,Vitevitch1997} and perceiving \citep{Dupoux1999,Kabak2007a,Massaro1983} certain types of nonce words judged to be phonotactically illicit.
For instance, \citet{GallagherInPress} finds that speakers of Quechua have difficulty repeating nonce words with multiple ejectives (e.g., [k'ap'i]), which do not occur in the language; this suggests that 
multiple ejective words are not merely absent, but actually difficult for these speakers to perceive, represent, or produce.

In English, sequences of adjacent obstruents which do not also agree in voice (e.g., \emph{a}[b.s]\emph{inth}) are quite rare within a word, and therefore a hetero-voiced obstruent cluster is a clue to the presence of a word boundary in running speech.
Infants \citep[e.g.,][]{Mattys2001b} and \citep{McQueen1998b,Norris1997} are both thought to use this heuristic for word recognition in experimental settings.\footnote{
    There is reason to distinguish the computations involved in using this word segmentation heuristic from those implicated by other phonotactic behaviors.
    Whatever the locus of [bnɪk]'s ill-formedness, for example, no segmentation into multiple words or morphs renders it a well-formed sequence.
    See \S\ref{ss:bh} for further discussion.}

\citet{Berent2001b} and \citet{Coetzee2008b} claim that non-word recognition latencies in lexical decision tasks reflect speakers' phonotactic knowledge, the hypothesis being that a phonotactically illicit nonce word will be rejected more quickly than a well-formed nonce word.
However, phonotactic constraints are often confounded with independent predictors of lexical decision latencies.
For instance, \citet{Coetzee2008b} finds that English speakers recognize [sp\ldots{}p] and [sk\ldots{}k] nonce monosyllables faster than [st\ldots{}t] nonce monosyllables in an auditory lexical decision task.
\citeauthor{Coetzee2008b} attributes this to ad hoc phonotactic constraints against the former sequences, but another explanation is available.
Even at an early stage of recognition, [stVt] is distinguished from [spVp, skVk] by its higher \emph{cohort density}: there are far more English words starting with initial [st] than with [sp] or [sk].
High cohort density is known to inhibit auditory processing of non-words \citep[e.g.,][]{Marslen-Wilson1978} and this alone could account for the processing difference.

\subsection{Loanword adaptation}

Loanword adaptation may provide further evidence for the psychological reality of phonotactic knowledge.
In Desano \citep{Kaye1974}, for instance, all underlying representations (URs) are either totally oral (e.g., [yaha] `to hear') or totally nasal (e.g., [ñãhã] `to enter'), and loanwords are made to conform to this generalization: Portuguese \emph{martelo} `hammer' is adapted as [barateru] and Spanish \emph{naranja} `orange' as [nãnãnã].
Presumably, some component of the synchronic grammar is responsible for the fact that the restriction over native vocabulary is extended to loanwords.

There are many other cases, however, where it is claimed that phonotactic restrictions are not extended to loanwords (e.g., \citealt{Clements1982}, \citealt[75]{Davidson1997}, \citealt[95]{Shibatani1973}, \citealt{Ussishkin2003}, \citealt{Vogt1954}; additional examples can be found throughout this dissertation).
Given the limited understanding of loanword adaptation at the present juncture, it may be premature to regard this inertness as strong evidence against the constraints in question, though it may be a useful diagnostic.

\subsection{Alternate phonologies}

Language games or speech disguises may also provide evidence for phonotactic knowledge \citep[e.g.,][]{Vaux2011}; an example appears in \S\ref{ss:bh} below.

\subsection{Lexical statistics}

Finally, phonotactic gaps or tendencies in the lexicon are often taken as evidence for phonotactic knowledge, under the hypothesis that grammatical constraints are the cause of these lexical generalizations.
Chapters \ref{turkish}--\ref{gaps} consider in detail the evidentiary status of these lexical tendencies.

\section{The grammatical basis of phonotactic knowledge}

@@

Early generative phonologists posited that phonotactic ill-formedness derives solely from \emph{morpheme structure constraints}, restrictions on underlying representations \citep{Chomsky1965,SPE,SPR,Halle1962}.
\citet{Stanley1967} distinguishes between two types.
\emph{Segment structure constraints} impose restrictions on the underlying segment inventory.
For example, voicing is non-contrastive for /ts, tʃ, x/ in Russian \cite[22]{SPR}: [dz, dʒ, ɣ] appear in surface, but not underlying, representations.
\emph{Sequence structure constraints} apply to underlying sequences; an example is given below.

@@

\noindent
This sequence structure constraint specifies the second of a sequence of word-initial consonants as a liquid, so as to preclude underlying */bnɪk/, for example.

However, \citet{Shibatani1973} shows quite convincingly that not all wordlikeness contrasts can be expressed as constraints on URs.
In German, for instance, obstruent voicing is contrastive, but neutralizes finally: e.g., [ɡʀaːt]-[ɡʀaːtə] `ridge(s)' vs.~[ɡʀaːt]-[ɡʀaːdə] `degree(s)'.
By hypothesis, the latter root ends in /d/, so the constraint against final voiced obstruents is specific to surface representations.
Yet, \citeauthor{Shibatani1973} claims, German speakers judge voiced obstruent-final nonce words to be ill-formed.\footnote{
    Voicing of final obstruents is usually lost in German loanword adaptation: e.g., English \emph{hot dog} becomes [hɑt dɔk] \citep[506]{Ussishkin2003}.}

@@

Whereas \citeauthor{Shibatani1973} argues that morpheme structure constraints are insufficient to account for speakers' phonotactic knowledge, other authors observe the tendency for structural descriptions of phonological processes to reappear among the morpheme structure constraints on the same language (e.g., \citealt{Hale1965}:297, \citealt{Kisseberth1970b}, \citealt{Postal1968}:212f.).
Russian obstruent voice assimilation, discussed above, provides an example of this type: there are no hetero-voiced obstruent clusters in either underlying or surface representations.
These two facts are tantalizingly similar, but are treated as separate if a distinction between morpheme structure constraints and phonological processes is drawn.
This is a special case of what \citet{Kisseberth1970b} calls \emph{conspiracies}, which \citet[136]{KK77} term the \emph{duplication problem}.
\citet[205f.]{Dell1973} and \citet[28f.]{Stampe1973} argue that the distinction between constraints on URs and alternations is artificial, and that these different levels of description are related by a principle now known as \emph{Stampean occultation} \citep[54]{OT}.
In a language like Russian, in which surface obstruent clusters exceptionlessly agree in voicing, there is simply no reason for the language acquisition device to posit underlying hetero-voiced obstruent clusters: obstruent voice assimilation ``occults'' underlying */kb/, for instance. Were such an underlying form posited, it would surface as [gb] in all contexts.\footnote{
    \emph{Lexicon Optimization} \citep[209]{OT} implements a form of Stampean occultation notable in that it projects all non-alternating surface segments directly into URs.
    For instance, in English, Lexicon Optimization demands underlying /ŋ/ in words like \emph{bank}, though [ŋ] could be otherwise be analyzed as an allophone of /n/ before velar consonants (e.g., \citealt[65f.]{Borowsky1986}, \citealt[85]{SPE}, \citealt[62]{Halle1985a}), eliminating /ŋ/ from the phoneme inventory.
    However, this is not core to \citeauthor{Dell1973} and \citeauthor{Stampe1973}'s insight about the relationship between surface and underlying sequence structure restrictions.
    For instance, the hypothetical \{bæŋk\} posited by Lexicon Optimization could be revised to /bænk/, and take a \emph{free ride} (in the sense of \citealt{Zwicky1970}) on the process of nasal place assimilation found elsewhere in English (see \S\ref{npa}).
    Indeed, this seems desirable, since Lexicon Optimization forces a duplication between underlying and surface constraints.
%this point is made by \citet{Chomsky1957} and \citet[69f.]{Harris1960} with respect to the principle of biuniqueness, which enforces a similar restriction.
    For instance, [ŋ] does not appear word-initially and English speakers have considerable difficulty producing it in this position \citep{Rusaw2009}.
    The allophonic analysis of [ŋ] predicts this fact, since there is no way to derive the [ŋ] allophone in onset position.
    Future work will consider the how purely allophonic relationships might be acquired.}
    %Assuming Lexicon Optimization, the only alternative is to stipulate an ad hoc constraint against onset [ŋ] \citep[39]{Jusczyk2002}.}
In chapter \ref{gaps}, it is argued that constraints described in terms of non-contrastive prosodic structures can also be derived by Stampean occultation.
In an architecture like Lexical Phonology, it is even possible to apply a process to individual underlying representations (i.e., at the ``lexical level'') to enforce constraints which are not surface-true.
Consequently, morpheme structure constraints are otiose, their empirical converage entirely subsumed by phonological processes.

This argument against the distinction between morpheme structure constraints and phonological processes is quite similar to the famous argument (\citeyear{SPR}) against the morphophonemic/phonemic distinction.
The principle of biuniquness in vogue at that time separates neutralizing (morphophonemic) and non-neutralizing (phonemic) processes.
In Russian, obstruents participate in voice assimilation whether this neutralizes a phonemic distinction (\ref{rovs}a) or not (\ref{rovs}b): recall that there is no underlying /dʒ/ in Russian.
Under biuniqueness, there must be separate neutralizing and non-neutralizing variants of this process, however.
From this, \citeauthor{SPR} argues that biuniqueness (and the distinction between the morphophonemic and phonemic levels that follows from it) entails ``a significant increase in the complexity of the representation'' (24).
While \citet{Anderson2000} sketches an analysis which preserves biuniqueness without morphophonemic/phonemic duplication (see also \citealt{Kiparsky1985}), this requires further contested assumptions---contrastive underspecification (against which, see \citealt{Steriade1995}) and a Duke-of-York derivation.
This does not seem inconsistent with \citeauthor{SPR}'s claim that biuniquness imposes unnecessary additional complexities: under \citeauthor{Anderson2000}'s analysis the complexity is not a duplication, but rather a dependency on contentious theoretical assumptions.

\subsection{Static constraints}

Under the assumptions of Stampean occultation, there is no way to express substantive constraints on underlying representations except via phonological process.
In lexical phonology, such constraints are the consequences of phonological processes operating at the ``lexical'' level, and in an Optimality Theory grammar, these are high ranked markedness constraints.\footnote{In Optimality Theory, it is often assumed that all markedness constraints are ranked above all faithfulness constraints in the ``initial state'' of the learner \citep[e.g.,][]{Smolensky1996a}.
    This admits the possibility that markedness constraints not implicated by alternations will remain undominated \citep[e.g.,][]{Coetzee2008b}.
    It is difficult to evaluate this proposal in the absence of any substantive proposal for the contents of \textsc{Con}, the universal constraint sets, but this has the potential to blur the traditional distinction between accidental and structural phonotactic gaps: any exceptionless gap which corresponds to a markedness constraint \emph{in any language} will be accorded structural, inviolable status, a quite powerful prediction.
    If \textsc{Con} is sufficiently elaborate to incorporate constraints like the *[spVp] proposed by \citet{Coetzee2008b}, it might also incorporate a constraint like *[s.w], perhaps as a subcomponent of the so-called \emph{syllable contact law} \citep[e.g.,][]{Gouskova2004,Murray1983}, which disfavors syllable contact clusters with increasing sonority 
    The *[s.w] constraint is without exception in English, yet nonce words like \emph{teeswa} [tis.wa] seem quite unobjectionable.
    Numerous other examples of this type could be adduced.
    Consequently, this specific model is rejected here.}
A corrolary is that if some phonotactic constraint is not (or cannot be) identified with a phonological process, it does not exist in the minds of speakers at all.

This principle---\emph{no static phonotactic constraints}---has interesting ramifications for evaluating certain competing phonological analyses.
Consider Sanskrit aspiration alternations such as \emph{bodhati}-\emph{bhotsyati} `he wakes-he will wake'.
According to one analysis, which has precedents as far back as the grammar of Pāṇini, the root /budh/ undergoes a process shifting aspiration leftward in certain contexts (e.g., \citealt{Borowsky1983}, \citealt{Hoenigswald1965}, \citealt{Kaye1985}, \citealt{Sag1974}, \citeyear{Sag1976}, \citealt{Schindler1976}, \citealt{Stemberger1980}, \citealt{Whitney1889}:\S141f.).
An alternative analysis posits an underlying /bhudh/ and a process of aspirate dissimilation, a synchronic analogue of Grassman's Law (e.g., \citealt{Anderson1970}, \citealt{Hoard1975}, \citealt{Kiparsky1965}:\S3.2, \citealt{Phelps1973}, \citealt{Phelps1975b}, \citealt{Zwicky1965}:109f.).
Under the latter analysis, multiple surface aspirates (e.g., hypothetical *\emph{bhodhati}) are phonotactically marked; the former account makes no such prediction.
The principle of no static phonotactic constraints, if it can be maintained, could in theory adjudicate between these competing analyses: if the postulated constraint on multiple surface aspirates is active in wordlikeness judgements or loanword adaptation, for instance, this rules out the former account.

\subsection{Acquisition and the (in)dependence of phonotactic knowledge}
\label{ss:aerbpp}

\citet{Hayes2004b} argues that phonotactic learning occurs before lexical or phonological acquisition, and is therefore must be independent of other types of grammatical knowledge.
This is not fully consistent with the available experimental evidence available, however.

Typically-developing infants know their names and the names of their caretakers as early as 4 months of age \citep{Bortfeld2005,Mandel1995,Tincoff1999}.
While infants are born able to distinguish between monosyllabic and bisyllabic words, the youngest infants are thought to recognize syllables ``holistically'' rather than as segment sequences \citep{Bertoncini1981,Eimas1999,Jusczyk1987}.
As early as 6 months of age, infants know the visual referents of familiar words presented auditorily \citep{Bergelson2012}.
At 7.5 months, phonological representations are sufficiently detailed to allow infants to discriminate between words like \emph{cup} and mispronunciations like \emph{*tup} \citep{Jusczyk1995}.
By 8 months, infants are able to locate both familiar and novel words in continuous speech \citep{Jusczyk1997,Seidl2006}.

While very few studies have investigated young infants' knowledge of phonological alternations, this is the subject of a fascinating study by \citet{White2008}.
Simplifying somewhat, the experimenters expose 8.5-month-old infants to an artificial language in which fricative voicing is contrastive, but voiced and voiceless variants of plosives are in complementary distribution, appearing only after vowels (\emph{na-bevi}) and after voiceless consonants (\emph{rot-pevi}), respectively.
After familiarization, infants prefer to listen to nonce words preserving this complementary distribution of stops over nonce words which disrupt this distribution (e.g., \emph{na-poli}, \emph{rot-boli}), suggesting that the infants have extracted an allophonic generalization concerning plosive voicing.

Crucially, all of this occurs \emph{before} there is any evidence for phonotactic knowledge, for which the earliest evidence begins at approximately 9 months of age \citep{Friederici1993,Jusczyk1993b,Jusczyk1994}.
While there are many gaps in current understanding (and keeping in mind the usual caveats about a strong interpretation of negative results), there is no reason to think that phonotactic knowledge is acquired before non-trivial amounts of lexical and phonological acquisition.
If anything, infants's phonotactic knowledge appears to come online at approximately the same point as highly specific phonological representations including subsyllabic units and the ability to extract allophonic generalizations, as predicted by the firm link between phonological and phonotactic knowledge advocated for here.

@@

Chapter \ref{gradience} considers evidence from wordlikeness rating tasks. It is argued that intermediate well-formedness ratings are obtained whether or not the categories in question are graded. A primitive categorical model of wordlikeness using prosodic representations is outlined, and shown to predict English speakers' wordlikeness judgements as accurately as state-of-the-art gradient wellformedness models. Once categorical effects are controlled for, gradient models are uncorrelated with well-formedness ratings.

Chapter \ref{turkish} considers the relationship between lexical generalizations, phonological alternations, and speakers' nonce word judgements with a focus on Turkish vowel patterns. It is shown that even exception-filled phonological generalizations have a robust effect on wellformedness judgements, but that statistically reliable phonotactic generalizations go unlearned when they are not derived from phonological alternations.

Chapter \ref{gaps} investigates the role of phonological alternations in constraining lexical entries, focusing specifically on medial consonant clusters in English. Static phonotactic constraints previously proposed to describe gaps in the inventory of medial clusters are shown to be statistically unsound, whereas phonological alternations impose robust restrictions on the cluster inventory. The remaining gaps in the cluster inventory are attributed to the sparse nature of the lexicon, not static phonotactic restrictions.

@@

%\begin{example}[German \emph{ich}- and \emph{ach-laut}]
%\begin{tabular}{l l l}
%a. & [buːx]   & `book'           \\
%   & [tɔxtər] & `daughter'       \\
%   & [naxt]   & `night'          \\
%b. & [siçt]   & `view'           \\
%   & [ʃpeçt]  & `woodpecker'     \\
%   & [ɡərʏçt] & `rumor'          \\
%   & [knøçəl] & `ankle, knuckle' \\
%   & [flɛçə]  & `surface'        \\
%\end{tabular}
%\end{example}
%
%\emph{Umlaut}, the fronting (and raising) of back vowels in certain morphological contexts, produces the front variant of the dorsal fricative; e.g., [lɔx]-[løçər] `hole-holes', \emph{B}[uːx]-\emph{B}[yːç]\emph{er} `book-books',
% 
%Throughout the term \emph{lexicon} is used in a specific sense of the set underlying representations in some language; this is not meant to imply a position on the possibility that larger, composite linguistic representations are also stored in lexical memory (thought see \citealt{LignosInPressa} for some recent experimental evidence bearing on this question).
%I reject the possibility that phonotactics do exist altogether.
